---
title: "Homework7"
author: "Eurydice Aboagye"
date: "2023-03-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Simulating and Fitting Data Distributions  



```{r}
z <- read.table("BiofilmCounts.csv",header=TRUE,sep=",") ##BiofilmCounts is my own generated data
z <- z[complete.cases(z), ]
str(z)
summary(z)

```

```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

```{r}
p1 <- ggplot(data=z, aes(x=logcfu, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1) # a histogram on that shows the distribution of my primary outcome, log cfu/coupon
```

```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1) #Empirical density curve
```

```{r}
normPars <- fitdistr(z$logcfu,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] ## fitting a normal distribution to my data and grabing the maximum likelihood estimators of the two parameters of the normal: The mean and the variance


```
That generated a list of 5:"n", "loglik", "estimate", "mean" and "sd", all pf which are the parameters of the normal distribution 

I then passed these parameters to the stat_function as ff:

```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$logcfu),len=length(z$logcfu))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$logcfu), args = list(mean = meanML, sd = sdML))
 p1 + stat
 
 #this generated a probability density for the normal distribution
```


I then added in the curve for the exponential as ff:

```{r}
expoPars <- fitdistr(z$logcfu,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$logcfu), args = list(rate=rateML))
 p1 + stat + stat2
 
```

I then plotted the uniform probability density which did not require the fitdisr function, because the max likelihood estimators in this case, are just the min and max of the data


```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$logcfu), args = list(min=min(z$logcfu), max=max(z$logcfu)))
 p1 + stat + stat2 + stat3

```

I then plotted the  gamma probability density:

```{r}
gammaPars <- fitdistr(z$logcfu,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$logcfu), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4

```

The beta probability was shown on its own plot, because the original data was re-scaled to values between 0 and 1.

```{r}
pSpecial <- ggplot(data=z, aes(x=logcfu/(max(logcfu + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$logcfu/max(z$logcfu + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$logcfu), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```

The gamma probability density appears to be my best fit.

I then simulated a new data set, with the same length as the log cfu vector 


```{r}
y <- rgamma(n=length(z$logcfu), shape =shapeML, rate=rateML)

y <- data.frame(1:length(z$logcfu),y)
names(y) <- list("ID", "new_cfu")
str(y)
summary(y)
psim<- ggplot(data=y, aes(x=new_cfu),  y=..density..) + 
  geom_histogram(color="grey60", fill="cornsilk", size=0.2)
print(psim)

##now for the gamma distribution curve...

## first, i determine the parameters..

simgammaPars<-fitdistr(y$new_cfu, "gamma")
print(simgammaPars)
 psim2<- psim + geom_density(linetype="dotted",size=0.75)
print(psim2)


```

Now comparing to the original data set....

```{r}

p1 <- ggplot(data=z, aes(x=logcfu, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 

p1 <-p1 + geom_density(linetype="dotted",size=0.75)

print (p1)
```

the simulated data appears comparable to the original data
