---
title: 'Homework10: For Loops and Randomization Tests'
author: "Eurydice Aboagye"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question One  
Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output

```{r}
zero_counts <- function(x) {
  counter <- 0  # setting up the counter variable
  
  for (i in x) {
    if (i == 0) {      
      counter <- counter + 1  # adding 1 to counter each time there is a zero in the vector
    }
  }
  
  return(counter)  # Return the final count of zeroes
}


# Testing the function..

d<- c(rep(-1:3, times=5))
print(d)
zero_counts(x=d)

```

### Question Two  
Use subsetting instead of a loop to rewrite the function as a single line of code.

```{r}
zero_counter <- function(x){
  return(sum(x==0))# return the sum of elements in x that are equal to zero
}

#  Testing zero counter
zero_counter(x=d)

```

### Question Three  
Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
#######################################
# FUNCTION: matrix_builder
# packages: none
# purpose:Building a dummy matrix 
# input: two integers, x and y, representing the number of rows and column in a matrix
# output: a matrix of specified dimensions in which each element is the product of the row number x the column number y
# -------------------------------------

matrix_builder <- function(x=NULL, y=NULL) {
  if(is.null(x)) m <- matrix(data = d, nrow = 5, ncol = 5, byrow = TRUE) # setting a null to mostly to check if function works
  else{
  m <- matrix(data=outer(1:x, 1:y, "*"), # found this outer function online. It's output is a matrix that applies a specified function to elements, in this case, a multiplier
                nrow=x,
                ncol=y,
                byrow=FALSE)}
  return(m)
  }

# Testing matrix builder

matrix_builder()

matrix_builder(6,10)

```


### Question Four  

a) Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.

```{r}
# create treatment groups
trt_group <- c(rep("Control",4),rep("Treatment1",5), rep("Treatment2",5))

# create response variable
res <- c(runif(4) + 1, runif(5) + 10, runif(5) + 6)
print(res)
print(trt_group)

# combine into a data frame
df<- data.frame(group= trt_group, response=res)
print(df)

```

b) Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data.

```{r}
remix_var<- function(x){
  .<- data.frame(group=x[,1], response=sample(x[,2]))
  df1<- aggregate(.[,2], list(.[,1]), FUN=mean) 
  return(df1)
}

#testing remix_var

remix_var(x=df)

```

c) Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.

```{r}
# creating an empty dataframe to hold final output

loopdf<- data.frame(RepNumber=1:100, ControlMeans= rep(NA, 100), T1Means= rep(NA, 100), T2Means= rep(NA, 100))

head(loopdf)

s<-seq(1:100) # created variable to hold number of times to carry out loo operation

for(i in seq_along(s)){ 
  . <- remix_var(x=df) 
  .<- list(ControlMeans=.[1,2],
           T1Means=.[2,2],
           T2Means=.[3,2])
  loopdf[i,2:4]<- unlist(.) #unlist and coy into last 3 columns of empty data frame
}

print(loopdf)

```

d) Use qplot() to create a histogram of the means for each reshuffled group. 

```{r}

library(ggplot2)
qplot(x=loopdf$ControlMeans, geom = "histogram", xlab = "Mean", ylab = "Replicates", main = "Means of Control Group")

qplot(x=loopdf$T1Means, geom = "histogram", xlab = "Mean", ylab = "Replicates", main = "Means of Treatment1 Group")

qplot(x=loopdf$T2Means, geom = "histogram", xlab = "Mean", ylab = "Replicates", main = "Means of Treatment2 Group")

```
